---
title: "CaseStudy_01"
author: "Stephen Johnson"
date: "February 15, 2019"
output: 
  html_document:
  keep md: true
---

## Import the files "beer" and "breweries" into R
```{R echo=TRUE}
setwd("C:/Users/sj103/OneDrive/Documents/MSDS_Working/CaseStudy_01")
beers <- as.data.frame(read.csv(file="Beers.csv", header=TRUE, sep=","))
breweries <- as.data.frame(read.csv(file="Breweries.csv", header=TRUE, sep=","))
```

## 1. How many breweries are present in each state?
```{R echo=TRUE}
library(dplyr)
library(knitr)
counts <- breweries %>% # %>% pipe function
  group_by(State) %>%
  summarize(count =n()) %>%
  arrange(desc(count))
counts2 <- cbind(counts[1:25, ], counts[26:50, ])
kable(counts2)
kable(head(counts, 5))
kable(tail(counts, 5))
```


##2. Merge beer data with the breweries data. Print the 1st 6 and last 6
```{R echo=TRUE}
colnames(beers) <- c("Beer", "Beer_ID", "ABV", "IBU", "Brewery_id", "Style", "Ounces")
colnames(breweries) <- c("Brewery_id", "Brewery", "City", "State")
final <- merge(beers, breweries, by = "Brewery_id")
kable(head(final, 6))
kable(tail(final, 6))
```

##3. Report the number of NA's in each column
```{R echo=TRUE}
sapply(final, function(x) sum(is.na(x)))
summary(final)
```

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{R echo=TRUE}
median_vals <- final %>%
  group_by(State) %>% 
  summarise(median_abv = median(ABV, na.rm = T),
            median_ibu = median(IBU, na.rm = T),
            mean_abv = mean(ABV,na.rm =T))

library(ggplot2)
ggplot(median_vals, aes(x = reorder(State, -median_abv), y = median_abv)) + 
  geom_bar(stat = 'identity', color = 'black', fill = 'yellow2') + 
  labs(x = 'State', y = "Median ABV", title = "Median ABV by State") + 
  theme(axis.text.x = element_text(size = 10, angle = 90))
ggplot(median_vals, aes(x = reorder(State, -median_ibu), y = median_ibu)) +
  geom_bar(stat = 'identity', color = 'black', fill = 'darkorange1') +
  labs(x = 'State', y = "Median IBU", title = "Median IBU by State") +
  theme(axis.text.x = element_text(angle = 90))
  ```
```

### 5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
max_vals = final %>% 
  group_by(State) %>% 
  summarise(max_ibu = max(IBU, na.rm = T),
            max_abv = max(ABV, na.rm = T))

# group_by, summmarise, arrange, slice, select are all from dplyr. 
max_vals %>% arrange(-max_ibu) %>% slice(1)
max_vals %>% arrange(-max_abv) %>% slice(1)

```

### 6.	Summary statistics for the ABV variable.
```{r}
final %>% select(ABV) %>% summary()
summary(final)
```

### 7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r}
ggplot(final, aes(x = IBU, y = ABV)) +
  geom_point(alpha = .8, shape = 1, color = 'darkorange') + 
  geom_smooth(method= 'lm')

```
