---
title: "CaseStudy_01"
author: "Stephen Johnson"
date: "February 15, 2019"
output: 
  html_document:
  keep md: true
---

### Import the files "beer" and "breweries" into R
```{r echo=TRUE}
setwd("C:/Users/sj103/OneDrive/Documents/MSDS_Working/CaseStudy_01")
beers <- as.data.frame(read.csv(file="Beers.csv", header=TRUE, sep=","))
breweries <- as.data.frame(read.csv(file="Breweries.csv", header=TRUE, sep=","))
state_pop <- as.data.frame(read.csv(file="StatePop.csv", header=TRUE, sep=","))
```

##1. How many breweries are present in each state?
```{R echo=TRUE}
library(dplyr)
library(knitr)
counts <- breweries %>%
  group_by(State) %>%
  summarize(count =n()) %>%
  arrange(desc(count))
counts2 <- cbind(counts[1:25, ], counts[26:50, ])
kable(counts2)
kable(head(counts, 5))
kable(tail(counts, 5))
```

### Mapping - Breweries per State
```{r echo=TRUE}
library(usmap)
library(ggplot2)
breweries_per_state = counts %>% 
  rename(state = State) %>% 
  mutate(state = gsub(" ","",state))

plot_usmap(data = breweries_per_state, values = "count")+
scale_fill_continuous(low = "white", high = "red")
```

### Breweries per Capita/Mapping
```{r echo=TRUE}
state_info <- data.frame(state = state.name, state_abb = state.abb)
state_info <- state_info %>% 
  merge(state_pop %>% rename(state = STATE), by = 'state') 
breweries_updated <- breweries_per_state %>% 
  merge(state_info, by.x = "state", by.y = "state_abb", all.x = TRUE) %>% 
  mutate(population = as.numeric(gsub(",", "", as.character(POPULATION))),
  bwr_per_capita = count/population) %>%
  arrange(desc(bwr_per_capita))
bwr_pc <- cbind(breweries_updated$state, breweries_updated$bwr_per_capita)
colnames(bwr_pc) <- c("state", "pcap")
kable(head(bwr_pc))

plot_usmap(data = breweries_updated, values = "bwr_per_capita")+
  scale_fill_continuous(low = "white", high = "red")
```

##2. Merge beer data with the breweries data. Print the 1st 6 and last 6
```{R echo=TRUE}
colnames(beers) <- c("Beer", "Beer_ID", "ABV", "IBU", "Brewery_id", "Style", "Ounces")
colnames(breweries) <- c("Brewery_id", "Brewery", "City", "State")
final <- merge(beers, breweries, by = "Brewery_id")
kable(head(final, 6))
kable(tail(final, 6))
```

##3. Report the number of NA's in each column
```{R echo=TRUE}
sapply(final, function(x) sum(is.na(x)))
summary(final)
```

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{R echo=TRUE}
median_vals <- final %>%
  group_by(State) %>% 
  summarise(median_abv = median(ABV, na.rm = T),
            median_ibu = median(IBU, na.rm = T))
median_vals2 <- median_vals %>%
   rename(state = State) %>%
   mutate(state = gsub(" ", "",state))
head(median_vals2)
library(ggplot2)
ggplot(median_vals2, aes(x = reorder(state, -median_abv), y = median_abv)) + 
  geom_bar(stat = 'identity', width = 0.9, position = position_nudge(x = 0.5),
  color = 'black', fill = 'yellow2') + 
  labs(x = 'State', y = "Median ABV", title = "Median ABV by State") + 
  theme(axis.text.x = element_text(size = 10, angle = 90))
ggplot(median_vals2, aes(x = reorder(state, -median_ibu), y = median_ibu)) +
  geom_bar(stat = 'identity', width = 0.9, position = position_nudge(x = 0.5),
  color = 'black', fill = 'darkorange1') +
  labs(x = 'State', y = "Median IBU", title = "Median IBU by State") +
  theme(axis.text.x = element_text(angle = 90))
```

##5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r echo=TRUE}
med_abv <- median_vals2 %>%
  arrange(desc(median_abv))
kable(head(med_abv))
med_ibu <- median_vals2 %>%
  arrange(desc(median_ibu))
kable(head(med_ibu))
max_vals = final %>% 
  group_by(State) %>% 
  summarise(max_ibu = max(IBU, na.rm = T),
            max_abv = max(ABV, na.rm = T))
max_vals %>% arrange(-max_ibu) %>% slice(1)
max_vals %>% arrange(-max_abv) %>% slice(1)
```

##6.	Summary statistics for the ABV variable.
```{r echo=TRUE}
final %>% select(ABV) %>% summary()
summary(final)
```

##7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r echo=TRUE}
ggplot(final, aes(x = IBU, y = ABV)) +
  geom_point(alpha = .8, shape = 1, color = 'darkorange') + 
  geom_smooth(method= 'lm')
```
